---
title: "STA 325 Case Study"
authors: "Matthew Cui, Phillip Harmadi, Glen Morgenstern, and Joe Wang"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and data
```{r echo=FALSE, message=FALSE}
library(tidyverse)
train <- read_csv("data-train.csv")
test <- read_csv("data-test.csv")
train
cor(train)
test
```

## Exploratory Data Analysis

```{r eda}
# We transform the 'Fr' variable using the sigmoid function so that this variable 
# will be within a finite range.

train1 <- train %>%
  mutate(Fr_sigmoid = 1 / ( 1 + exp(-Fr))) %>%
  subset(select = c(1:3, 8, 4:7))

train1
cor(train1)

test1 <- test %>%
  mutate(Fr_sigmoid = 1 / ( 1 + exp(-Fr))) %>%
  subset(select = c(1:3, 8, 4:7))

# R_moment_2 is almost perfectly correlated with R_moment_3 and R_moment 4.
```

We will try to create these 4 models:

- **Response**: R_moment_1 & **Predictors (Main Effects)**: St, Re, Fr_sigmoid

We will attempt to use a combination of subset selection, polynomial, 
transformation, and interaction variables.

- **Response**: R_moment_2 & **Predictors (Main Effects)**: St, Re, Fr_sigmoid, R_moment_1

We will attempt to use a combination of subset selection, polynomial, 
transformation, and interaction variables. We will also include R_moment_1 since
it is has significant positive relationship with R_moment_2 (~0.63).

- **Response**: R_moment_3 & **Predictors (Main Effects)**: R_moment_2

We know that R_moment_2 is almost perfectly correlated (>0.99) with R_moment_3, 
so only using one predictor variable is enough. We try to avoid overfitting by
using only R_moment_2 as our only predictor to predict R_moment_3. We will
attempt to use polynomial and transformation variables.

- **Response**: R_moment_4 & **Predictors (Main Effects)**: R_moment_2, R_moment_3

Same reasoning - R_moment_2 and R_moment_3 are almost perfectly correlated
with R_moment_4. We will only use these 2 predictors and will attempt to use
both transformation and interaction variables (since R_moment_2 and R_moment_3
are also highly correlated to each other).

## Predictive models

```{r predictive-models}

```

## Apply to test data

```{r test-data}

```

